---
title: "Data Wrangling I"
author: Ekaterina Hofrenning
output: github_document
---
Load packages

```{r}
library(tidyverse)
```

Import `FAS_litters.csv` using a relative path:

```{r}
litters_df = read_csv("data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)


```

Import `FAS_litters.csv` using an absolute path.

```{r}
litters_df = read_csv("~/Documents/P8105/data_wrangling_i/data/FAS_litters.csv")
```

Using relative paths is better long-term because it ensures that everything is loaded in a contained folder/R project


Import pups dataset:

```{r}
pups_df = read_csv("data/FAS_pups.csv")
pups_df = read_csv("~/Documents/P8105/data_wrangling_i/data/FAS_pups.csv")
```


```{r}
skimr::skim(litters_df)
```

Parsing columns

```{r}
litters_data = read_csv(file = "./data/FAS_litters.csv",
  col_types = cols(
    Group = col_character(),
    `Litter Number` = col_character(),
    `GD0 weight` = col_double(),
    `GD18 weight` = col_double(),
    `GD of Birth` = col_integer(),
    `Pups born alive` = col_integer(),
    `Pups dead @ birth` = col_integer(),
    `Pups survive` = col_integer()
  )
)

litters_data = read_csv(file = "./data/FAS_litters.csv",
  col_types = cols(
    Group = col_factor()
  )
)
```


Other File Formats:

Excel files
```{r}
library(readxl)
mlb11_data = read_excel("data/mlb11.xlsx", n_max = 20)
head(mlb11_data, 5)
```

SAS files
```{r}
library(haven)
pulse_data = read_sas("./data/public_pulse_data.sas7bdat")
head(pulse_data, 5)
```

Base R vs. ReadR: 

- `read_csv()` is better than `read.csv()`. read.csv() prints all observations while read_csv() shows less.

- write_csv() to the relative path to export data
